---
title: "clean_perception_analysis"
author: "Sille Hasselbalch Markussen"
date: "5/13/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load('tidyverse', 'lme4', 'lmerTest', 'ordinal', 'ggeffects', 'car', 'stringr', 'dplyr', 'scales')

```

# Data load and preprocessing
```{r message=FALSE}
# Define the path to your folder
folder_path <- "perception2904"

# Get a list of all CSV files in that folder
csv_files <- list.files(path = folder_path, pattern = "*.csv", full.names = TRUE)

# Read and combine them into one dataframe
df <- csv_files %>%
  map_dfr(read_csv, show_col_types = FALSE)
```

PReprocessing
```{r}
# remove test trial (first 3 trials for each participant)
df <- df %>%
  group_by(participantID) %>%
  mutate(row_in_group = row_number()) %>%
  filter(row_in_group > 3) %>%
  select(-row_in_group) %>%
  ungroup()
```

```{r}
df <- df %>%
  mutate(
    labelPosition = as.factor(labelPosition),
    labelType = as.factor(labelType),
    brand = as.factor(brand),
    participantID = as.factor(participantID),
    accuracy = if_else(brand == labelType, 1, 0),
    labelPosition = dplyr::recode(labelPosition,
                                "U" = "byLabels",
                                "D" = "byPrice")
  )
```

Identifying outliers
```{r}
# Log-transform Reaction Times
df <- df %>%
  mutate(log_rt = log(reactionTime))

# Calculate Log Mean and SD (for Trimming)
log_mean <- mean(df$log_rt, na.rm = TRUE)
log_sd <- sd(df$log_rt, na.rm = TRUE)

# Trim Trials Outside ±2.5 SD on Log Scale
df_trimmed <- df %>%
  filter(log_rt >= (log_mean - 2.5 * log_sd),
         log_rt <= (log_mean + 2.5 * log_sd))
```



# Models

## Reaction time

Label position
```{r}
df_trimmed$labelPosition <- relevel(df_trimmed$labelPosition, ref = "byLabels")
rt_position <- lmer(log_rt ~ labelPosition + (1 | participantID), data = df_trimmed)

rt_position_slope <- lmer(log_rt ~ labelPosition + (1 + labelPosition| participantID), data = df_trimmed)

anova(rt_position, rt_position_slope, test = "Chisq")
```

```{r}
summary(rt_position)
```

Number of other labels
```{r}
rt_nLabels <- lmer(log_rt ~ n_labels + (1 | participantID), data = df_trimmed)
rt_nLabels_slope <- lmer(log_rt ~ n_labels + (1 + n_labels | participantID), data = df_trimmed)

# testing for the better model
anova(rt_nLabels, rt_nLabels_slope, test = "Chisq")
```

```{r}
summary(rt_nLabels)
```

Placement * n labels
```{r}
rt_interaction <- lmer(log_rt ~ n_labels:labelPosition + (1 | participantID), data = df_trimmed)
summary(rt_interaction)
```


## Accuracy

Label position
```{r}
ac_position <- glmer(accuracy ~ labelPosition + (1 | participantID), data = df_trimmed, family = binomial) 
ac_position_slope <- glmer(accuracy ~ labelPosition + (1 + labelPosition | participantID), data = df_trimmed, family = binomial) 

# checking for the best model
anova(ac_position, ac_position_slope, test = "Chisq")
```

```{r}
summary(ac_position_slope)
```

Number of labels
```{r}
ac_nLabels <- glmer(accuracy ~ n_labels + (1 | participantID), data = df_trimmed, family = binomial) 
ac_nLabels_slope <- glmer(accuracy ~ n_labels + (1 + n_labels | participantID), data = df_trimmed, family = binomial) 

# testing for the best model
anova(ac_nLabels, ac_nLabels_slope, test = "Chisq")
```

```{r}
summary(ac_nLabels)
```

Placement * n labels
```{r}
ac_interaction <- lmer(accuracy ~ n_labels:labelPosition + (1 | participantID), data = df_trimmed)
summary(ac_interaction)
```


# Plots - REPORT

RT Number of labels
```{r}
plot_data <- ggpredict(rt_nLabels, terms = "n_labels")

plot_data$predicted_ms <- exp(plot_data$predicted)
plot_data$predicted_s <- plot_data$predicted_ms/1000

ggplot(plot_data, aes(x = x, y = predicted_s)) +
  geom_line(size = 1.5, color = "#095979") +  # Line in stakeholder blue
  annotate("text", x = 4, y = 3.790, label = "+ 3,1%", color = "#095979", 
           fontface = "bold", family = "Times", size = 4, hjust = -0.2, vjust = 0.5) +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4, expand = c(0, 0)) +
  coord_cartesian(ylim = c(3.5, 4), clip = "off") +
  scale_y_continuous(labels = comma_format(decimal.mark = ","))+
  labs(
    x = "Antal andre mærker",
    y = "Svartid (sekunder)"
  ) +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(color = "#095979"),    
    axis.ticks.y = element_line(color = "#095979"),  
    axis.ticks.x = element_blank(),                  
    axis.title = element_text(color = "#095979"),
    axis.text = element_text(color = "#095979"),
    plot.margin = margin(5.5, 60, 5.5, 5.5), 
    plot.title = element_blank()
  )

ggsave('rt_nLabels_plot.png', bg = "transparent", width = 6, height = 4, dpi = 300)
```

ACC Number of labels
```{r}
plot_acc_nlabels <- ggpredict(ac_nLabels, terms = "n_labels")

ggplot(plot_acc_nlabels, aes(x = x, y = predicted)) +
  geom_line(size = 1.5, color = "#095979") +  
  annotate("text", x = 4, y = 0.95, label = "- 2,1 pp", color = "#095979", 
           fontface = "bold", family = "Times", size = 4, hjust = -0.2, vjust = 0.5) +
  scale_x_continuous(limits = c(0, 4),breaks = 0:4, expand = c(0, 0)) +
  coord_cartesian( clip = "off") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0.8, 1)) +
  labs(
    x = "Antal andre mærker",
    y = "Andel af korrekte klimamærker"
  ) +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(color = "#095979"),    
    axis.ticks.y = element_line(color = "#095979"), 
    axis.ticks.x = element_blank(),                 
    axis.title = element_text( color = "#095979"),
    axis.text = element_text(color = "#095979"),
    plot.margin = margin(5.5, 60, 5.5, 5.5),  
    plot.title = element_blank()
  )


ggsave('acc_nLabels_plot.png', bg = "transparent", width = 6, height = 4, dpi = 300)
```

RT Interaction
```{r}
plot_data_interaction$predicted_s <- plot_data_interaction$predicted_ms/1000

ggplot(plot_data_interaction, aes(x = x, y = predicted_s, color = group)) +
  geom_line(size = 1.5) +
  scale_color_manual(
    values = c(
      "byLabels" = "#007A6C",   # teal
      "byPrice" = "#C29F36"     # gold
    )
  ) +
  labs(
    x = "Antal andre mærker",
    y = "Svartid (sekunder)"
  ) +
  # Percent annotations (slightly above end of each line)
  annotate("text", x = 4.05, y = 3.795, label = "+3,4 %", 
           color = "#007A6C", fontface = "bold", family = "Times", hjust = 0) +
  annotate("text", x = 4.05, y = 3.765, label = "+2,7 %", 
           color = "#C29F36", fontface = "bold", family = "Times", hjust = 0) +
  #annotate("text", x = 1.1, y = 3640, label = "18% lavere stigning*", 
  #       color = "#034D69", fontface = "italic", family = "Times", hjust = 0.5)+
  scale_x_continuous(breaks = 0:4, expand = c(0, 0)) +
  scale_y_continuous(labels = comma_format(decimal.mark = ","))+
  coord_cartesian(ylim = c(3.5, 4), clip = "off") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    legend.position = "none",
    plot.margin = margin(5.5, 80, 5.5, 5.5),  # top, right, bottom, left
    axis.title = element_text( color = "#034D69"),
    axis.text = element_text(color = "#034D69"),
    axis.line = element_line(color = "#034D69", linewidth = 0.6),  # add axis lines
    axis.ticks.y = element_line(color = "#034D69")                   # add tick marks
  )

ggsave("interaction_plot_rt.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")
```

ACC Interaction
```{r}
ggplot(plot_acc_interaction, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1.5) +
  scale_color_manual(
    values = c(
      "byLabels" = "#007A6C",   # teal 
      "byPrice" = "#C29F36"     # gold 
    )
  ) +
  labs(
    x = "Antal andre mærker",
    y = "Andel af korrekte klimamærker"
  ) +
  geom_text(
    data = subset(plot_acc_interaction, x == max(x)),
    aes(
      y = predicted - c(0.0098, 0.0108),
      label = c("Blandt mærker", "Ved prisen")
    ),
    color = c("#007A6C", "#C29F36"),
    hjust = -0.05,
    fontface = "bold",
    family = "Times"
  ) +
  annotate("text", x = 4.05, y = 0.89, label = "-6 pp", 
           color = "#007A6C", fontface = "bold", family = "Times", hjust = 0) +
  annotate("text", x = 4.05, y = 0.935, label = "-0,9 pp", 
           color = "#C29F36", fontface = "bold", family = "Times", hjust = 0) +
  scale_x_continuous(breaks = 0:4, expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0.8, 1)) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    legend.position = "none",
    plot.margin = margin(5.5, 80, 5.5, 5.5),  
    axis.title = element_text( color = "#034D69"),
    axis.text = element_text(color = "#034D69"),
    axis.line = element_line(color = "#034D69", linewidth = 0.6), 
    axis.ticks.y = element_line(color = "#034D69")                  
  )

ggsave('acc_interaction_plot.png', width = 6.5, height = 4, dpi = 300, bg = "transparent")
```


# Plots - THESIS

RT Placement
```{r}
# Create predicted data
plot_rt_position <- ggpredict(rt_position, terms = c( "labelPosition"))

# Back-transform to milliseconds
plot_rt_position$predicted <- exp(plot_rt_position$predicted)
plot_rt_position$conf.low  <- exp(plot_rt_position$conf.low)
plot_rt_position$conf.high <- exp(plot_rt_position$conf.high)


# Plot
ggplot(plot_rt_position, aes(x = x, y = predicted, color = x)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.1, position = position_dodge(0.5)) +
  scale_color_manual(
    values = c("byLabels" = "#344D69",
               "byPrice" = "#E6C24B"),
    labels = c("By other labels", "By the price")
  ) +
  labs(
    x = "",
    y = "Predicted Response Time (ms)",
    color = "Label position"
  ) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           
                              keywidth = unit(-5, "cm"),       
                              default.unit = "cm"))+
  coord_cartesian(ylim = c(3000, 4500)) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    axis.text.x = element_blank()
  )

ggsave("thesis_figures/rt_position_pointPlot.png", width = 5, height = 4, dpi = 300)
```

ACC Placement
```{r}
plot_acc_position <- ggpredict(ac_position_slope, terms = c( "labelPosition"))

ggplot(plot_acc_position, aes(x = x, y = predicted, color = x)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.1, position = position_dodge(0.5)) +
  scale_color_manual(
    values = c("byLabels" = "#344D69",
               "byPrice" = "#E6C24B"),
    labels = c("By other labels", "By the price")
  ) +
  labs(
    x = "",
    y = "Predicted Accuracy",
    color = "Label position"
  ) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           
                              keywidth = unit(-5, "cm"),         
                              default.unit = "cm"))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0.8, 1)) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    axis.text.x = element_blank()
  )

ggsave("thesis_figures/acc_position_pointPlot.png", width = 5, height = 4, dpi = 300)

```

RT n labels
```{r}
plot_data <- ggpredict(rt_nLabels, terms = "n_labels")

plot_data$predicted_ms <- exp(plot_data$predicted)
plot_data$conf.low_ms <- exp(plot_data$conf.low)
plot_data$conf.high_ms <- exp(plot_data$conf.high)

ggplot(plot_data, aes(x = x, y = predicted_ms)) +
  geom_line(color = "#84bf7c", size = 1.2) +
  geom_ribbon(aes(ymin = conf.low_ms, ymax = conf.high_ms), fill = "#84bf7c", alpha = 0.2) +
  coord_cartesian(ylim = c(3000, 4500)) +
  labs(
    x = "Number of Other Labels",
    y = "Predicted Reaxtion Time (ms)"
  ) +
  theme_minimal()

ggsave("thesis_figures/rt_nLabels_pointPlot.png", width = 5, height = 4, dpi = 300)
```

ACC n labels
```{r}
# Predictions
plot_acc_nlabels <- ggpredict(ac_nLabels, terms = "n_labels")

# Plot
ggplot(plot_acc_nlabels, aes(x = x, y = predicted)) +
  geom_line(color = "#8BC17A", size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "#8BC17A", alpha = 0.2) +
  labs(x = "Number of other labels", y = "Predicted accuracy") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0.8, 1)) +
  theme_minimal()

ggsave("thesis_figures/acc_nLabels_pointPlot.png", width = 5, height = 4, dpi = 300)
```


RT interaction
```{r}
plot_data_interaction <- ggpredict(rt_interaction, terms = c("n_labels", "labelPosition"))

plot_data_interaction$group <- factor(
  plot_data_interaction$group,
  levels = c("byLabels", "byPrice")
)

plot_data_interaction$predicted_ms <- exp(plot_data_interaction$predicted)
plot_data_interaction$conf.low_ms <- exp(plot_data_interaction$conf.low)
plot_data_interaction$conf.high_ms <- exp(plot_data_interaction$conf.high)

ggplot(plot_data_interaction, aes(x = x, y = predicted_ms, color = group, fill = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low_ms, ymax = conf.high_ms), alpha = 0.2, colour=NA) +
  scale_color_manual(
    values = c("byLabels" = "#034D69", "byPrice" = "#e4c358"),
    labels = c("byLabels" = "By other labels", "byPrice" = "By the price")
  ) +
  scale_fill_manual(
    values = c("byLabels" = "#034D69", "byPrice" = "#e4c358"),
    labels = c("byLabels" = "By other labels", "byPrice" = "By the price")
  ) +
  labs(
    x = "Number of Labels",
    y = "Predicted Reaction Time (ms)",
    color = "Label Position",
    fill = "Label Position"
  ) +
  coord_cartesian(ylim = c(3000, 4500), clip = "off") +
  theme_minimal()

ggsave('rt_interaction_plot.png', bg = "transparent", width = 6, height = 4, dpi = 300)
```

ACC interaction
```{r}
plot_acc_interaction <- ggpredict(ac_interaction, terms = c("n_labels", "labelPosition"))

plot_acc_interaction$group <- factor(
  plot_acc_interaction$group,
  levels = c("byPrice", "byLabels")
)

# Plot
ggplot(plot_acc_interaction, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  scale_color_manual(
    values = c("byLabels" = "#034D69", "byPrice" = "#e4c358"),
    labels = c("byLabels" = "By the other labels", "byPrice" = "By the price")
  ) +
  scale_fill_manual(
    values = c("byLabels" = "#034D69", "byPrice" = "#e4c358"),
    labels = c("byLabels" = "By the other labels", "byPrice" = "By the price")
  ) +
  labs(
    x = "Number of Labels",
    y = "Predicted Accuracy",
    color = "Label Position",
    fill = "Label Position"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0.8, 1)) +
  theme_minimal()

ggsave('acc_interaction_plot.png', bg = "transparent", width = 6, height = 4, dpi = 300)
```



# Assumptions

rt placement
```{r}
# Extract residuals and fitted values
resid_rt <- resid(rt_position)
fitted_rt <- fitted(rt_position)

# Histogram of residuals
hist(resid_rt, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_rt)
qqline(resid_rt)

# Residuals vs. fitted
plot(fitted_rt, resid_rt,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

rt n labels
```{r}
# Extract residuals and fitted values
resid_rt <- resid(rt_nLabels)
fitted_rt <- fitted(rt_nLabels)

# Histogram of residuals
hist(resid_rt, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_rt)
qqline(resid_rt)

# Residuals vs. fitted
plot(fitted_rt, resid_rt,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

rt interaction
```{r}
# Extract residuals and fitted values
resid_rt <- resid(rt_interaction)
fitted_rt <- fitted(rt_interaction)

# Histogram of residuals
hist(resid_rt, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_rt)
qqline(resid_rt)

# Residuals vs. fitted
plot(fitted_rt, resid_rt,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

acc placement
```{r}
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp^2)
  ratio <- Pearson.chisq / rdf
  p <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(chisq = Pearson.chisq, ratio = ratio, rdf = rdf, p = p)
}
```

```{r}
# overdispersion
overdisp_fun(ac_position_slope)

# predicted vs actual values
df_trimmed$predicted <- predict(ac_position_slope, type = "response")

ggplot(df_trimmed, aes(x = predicted, y = accuracy)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Predicted probability", y = "Observed accuracy")
```

acc n labels
```{r}
# overdispersion
overdisp_fun(ac_nLabels)

# predicted vs actual values
df_trimmed$predicted <- predict(ac_nLabels, type = "response")

ggplot(df_trimmed, aes(x = predicted, y = accuracy)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Predicted probability", y = "Observed accuracy")
```

acc interaction
```{r}
#overdispersion
overdisp_fun(ac_interaction)

# predicted vs actual values
df_trimmed$predicted <- predict(ac_interaction, type = "response")

ggplot(df_trimmed, aes(x = predicted, y = accuracy)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Predicted probability", y = "Observed accuracy")
```

