---
title: "clean_judgement_analysis"
author: "Sille Hasselbalch Markussen"
date: "5/13/2025"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
pacman::p_load('tidyverse', 'lme4', 'lmerTest', 'ordinal', 'ggeffects', 'car', 'stringr', 'emmeans', "ggrepel")
```


# Data load and preprocessing
```{r message=FALSE}
# Define the path to your folder
folder_path <- "judgement2804"

# Get a list of all CSV files in that folder
csv_files <- list.files(path = folder_path, pattern = "*.csv", full.names = TRUE)

# Read and combine them into one dataframe
df <- csv_files %>%
  map_dfr(read_csv, show_col_types = FALSE)

df <- df %>%
  filter(trialType != "test")
```

Rename variables, relevel and set order
```{r}
df <- df %>%
  rename(
    phase = trialType,
    condition = scaleCondition,
    trueLabel = basketLabel
  ) %>%
  mutate(condition = dplyr::recode(condition,
                            "labels" = "noSupport",
                            "original" = "semanticSupport",
                            "linear" = "visuospatialSupport"),
         phase =  dplyr::recode(phase,
                            "experimental" = "Intervention",
                            "baseline" = "Baseline")) 


df$phase <- as.factor(df$phase)
df$income <- as.factor(df$income)
df$occupation <- as.factor(df$occupation)
df$phase <- relevel(df$phase, ref = "Baseline")

# Set the correct order for condition
df$condition <- factor(df$condition, 
                       levels = c("noSupport", "semanticSupport", "visuospatialSupport"),
                       labels = c("No additional support", "Semantic support", "Visuospatial support"))
df$phase <- factor(df$phase, levels = c("Baseline", "Intervention"))

```

Error and accuracy
```{r}
# Convert labels to numeric as before
label_to_numeric <- function(label) {
  return(match(label, c("A", "B", "C", "D", "E")))
}

df$true_numeric <- label_to_numeric(df$trueLabel)
df$selected_numeric <- label_to_numeric(df$selectedLabel)

# Calculate error
df$error <- abs(df$selected_numeric - df$true_numeric)
df$signed_error <- df$selected_numeric - df$true_numeric

df <- df %>% 
  mutate(accuracy = as.numeric(selectedLabel == trueLabel))
```

High emission
```{r}
df <- df %>%
  mutate(high_impact = trueLabel %in% c("D", "E"))

df$high_impact <- factor(df$high_impact, levels = c(TRUE, FALSE), labels = c("High", "Low"))
df$high_impact <- factor(df$high_impact, levels = c("Low", "High"))
```

Mixed baskets
```{r}
df_items <- read_csv('items.csv',show_col_types = FALSE)

# Clean names in both dataframes (trim whitespace)
df_main <- df %>%
  mutate(itemNames = str_trim(itemNames)) %>%
  dplyr::mutate(row_id = dplyr::row_number())

df_items <- df_items %>%
  mutate(item = str_trim(item))  # Make sure 'item' matches itemNames

# Split the itemNames column so each row has one item
df_long <- df_main %>%
  separate_rows(itemNames, sep = ",") %>%
  mutate(itemNames = str_trim(itemNames))  # remove any whitespace

df_long <- df_long %>%
  left_join(df_items %>% select(item, label), by = c("itemNames" = "item"))

# Create basket type per basket (you need an ID column, here assumed as `row_id`)
df_basket_labels <- df_long %>%
  group_by(row_id) %>%
  summarise(
    labels_in_basket = list(unique(label)),
    is_mixed = any(label %in% c("D", "E")) & any(label %in% c("A", "B")),
    is_uniform_high = all(label %in% c("D", "E")),
    .groups = "drop"
)

# Merge back into original df
df_main <- df_main %>%
  mutate(row_id = row_number()) %>%
  left_join(df_basket_labels, by = "row_id")


# Create basket type factor
df_main <- df_main %>%
  mutate(basket_type = case_when(
    is_uniform_high ~ "uniform_high",
    is_mixed ~ "mixed",
    TRUE ~ "other"
  ))

df_main <- df_main %>%
  mutate(mixedBasket = basket_type == "mixed")

df_main$mixedBasket <- factor(df_main$mixedBasket, levels = c(TRUE, FALSE), labels = c("Mixed", "Other"))

df_main$mixedBasket <- factor(df_main$mixedBasket, levels = c("Other", "Mixed"))
```

Delta certainty
```{r}

# Compute mean certainty per participant per phase
certainty_summary <- df %>%
  group_by(participantId, phase) %>%
  summarise(mean_certainty = mean(certainty, na.rm = TRUE), .groups = "drop")

# Reshape to wide format
certainty_wide <- certainty_summary %>%
  pivot_wider(names_from = phase, values_from = mean_certainty)

# Calculate delta certainty
certainty_wide <- certainty_wide %>%
  mutate(delta_certainty = Intervention - Baseline)


# Prepare data for model - include condition
model_data <- df %>%
  select(participantId, condition) %>%
  distinct() %>%
  inner_join(certainty_wide, by = "participantId")
```




# Models

Absolute error
```{r}
absErrorModel <- lmer(error ~ phase*condition + (1 | participantId), data = df)
summary(absErrorModel)
```

```{r}
# Absolute error - clmm check
df$abs_error_factor <- as.factor(df$error)

clmm_error_model <- clmm(abs_error_factor ~ phase * condition + (1|participantId), data = df)
summary(clmm_error_model)
```

Directional error
```{r}
signedErrorModel <- lmer(signed_error ~ phase*condition + (1 | participantId), data = df)
summary(signedErrorModel)
```

High emission baskets - directional error
```{r}
model_highImpact <- lmer(signed_error ~ high_impact * phase * condition + (1 | participantId), data = df)
summary(model_highImpact)
```

Mixed baskets
```{r}
model_mixedBasket <- lmer(signed_error ~ mixedBasket * phase * condition + (1 | participantId), data = df_main)
summary(model_mixedBasket)
```

Certainty 
- Certainty ~ condition * phase
```{r}
certainty_model <- lmer(certainty ~ phase*condition + (1 | participantId),data = df)
summary(certainty_model)
```

- Certainty ~ Accuracy
```{r}
certaintyAccuracy_model <- lmer(certainty ~ accuracy*phase*condition + (1 | participantId),data = df)
summary(certaintyAccuracy_model)
```

- Delta certainty ~ condition
```{r}
# Run a linear model
delta_Certainty_model <- lm(delta_certainty ~ condition, data = model_data)
summary(delta_Certainty_model)
```



# Plots - THESIS
Manual colours:
```{r}
# Custom colours
farver <- c(
  "Visuospatial support" = "#e4c358",
  "Semantic support" = "#007A6C",
  "No additional support" = "#034D69"
)
```


Absolute error - point plot
```{r}
# Predictions
plot_abs_error <- ggpredict(absErrorModel, terms = c("phase", "condition"))

# Create shaded rectangle data for the Baseline phase
baseline_rect_abs <- data.frame(
  xmin = 0.5,
  xmax = 1.5,
  ymin = -Inf,
  ymax = Inf
)

ggplot(plot_abs_error, aes(x = x, y = predicted, color = group)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  geom_rect(data = baseline_rect_abs,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey60", alpha = 0.2, inherit.aes = FALSE) +
  scale_color_manual(values = farver, labels = c(
    "No additional support   ",         # add spaces after label
    "Semantic support   ",
    "Visuospatial support")) +
  labs(
    #title = "Accuracy (absolute error from correct label)",
    x = "",
    y = "Predicted Absolute Error",
    color = "Support condition"
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           # label right of dot
                              keywidth = unit(-5, "cm"),         # tiny space before label
                              default.unit = "cm"))+
  theme_minimal()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 8.5),
    legend.key = element_blank(),               
    legend.box.spacing = unit(-0.1, "cm"),
    plot.title = element_text(hjust = 0.5),
    plot.margin = margin(t = 5, r = 25, b = 5, l = 25)
  )+
  coord_cartesian(clip="off")




ggsave("thesis_figures/abs_error_pointPlot.png", width = 5, height = 4, dpi = 300)
```

Directional error - point plot
```{r}
# Predictions
plot_signed_error <- ggpredict(signedErrorModel, terms = c("phase", "condition"))

# Create shaded rectangle data for the Baseline phase
baseline_rect_signed <- data.frame(
  xmin = 0.5,
  xmax = 1.5,
  ymin = -Inf,
  ymax = Inf
)

ggplot(plot_signed_error, aes(x = x, y = predicted, color = group)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  geom_rect(data = baseline_rect_signed,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey60", alpha = 0.2, inherit.aes = FALSE) +
  scale_color_manual(values = farver, labels = c("No Additional support   ",         # add spaces after label
                                                 "Semantic support   ",
                                                 "Visuospatial support")) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           # label right of dot
                              keywidth = unit(-5, "cm"),         # tiny space before label
                              default.unit = "cm"))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  annotate("text", x = max(as.numeric(plot_signed_error$x)) + 0.18, 
           y = -0.05, label = "Perfect accuracy", hjust = 0, size = 2.6, color = "black") +  
  labs(
    #title = "General underestimation bias",
    x = "",
    y = "Predicted Directional Error",
    color = "Support condition"
  ) +
  theme_minimal()+
  scale_y_continuous(limits = c(-1, 1)) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 8.5),
        legend.key = element_blank(),
        legend.box.spacing = unit(-0.1, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.margin = margin(5, 25, 5, 25))+
  coord_cartesian(clip="off")


ggsave("thesis_figures/signed_error_pointPlot.png", width = 5, height = 4, dpi = 300, bg = "transparent")

```

High emission - point plot
```{r}


# Create label data manually for the right facet
label_data <- data.frame(x = 2.3, y = 0.08, label = "Perfect accuracy", facet = factor("High", levels = c("Low", "High")))

baseline_rect <- data.frame(
  facet = factor(c("Low", "High"), levels = c("Low", "High")),
  xmin = 0.5,
  xmax = 1.5,
  ymin = -Inf,
  ymax = Inf
)


# Predictions including high_impact
plot_highImpact_error <- ggpredict(model_highImpact, terms = c("phase", "condition", "high_impact"))


plot_highImpact_error$facet <- factor(plot_highImpact_error$facet, levels = c("Low", "High"))



# Plot
ggplot(plot_highImpact_error, aes(x = x, y = predicted, color = group)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  geom_rect(data = baseline_rect,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey60", alpha = 0.2, inherit.aes = FALSE) +
  facet_wrap(~facet, labeller = as_labeller(c("High" = "High-emission baskets", "Low" = "Low-emission baskets"))) +
  scale_color_manual(values = farver, labels = c("No Additional support   ",         # add spaces after label
                                                 "Semantic support   ",
                                                 "Visuospatial support")) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           # label right of dot
                              keywidth = unit(-5, "cm"),         # tiny space before label
                              default.unit = "cm"))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_label(data = label_data, aes(x = x, y = y, label = label),
             inherit.aes = FALSE, color = "black", fill = "white",
             label.size = NA, size = 2.6)+

  labs(#title = "High-Emission bias",
    x = "",
    y = "Predicted Directional Error",
    color = "Support condition"
  ) +
  scale_y_continuous(limits = c(-1.1, 1)) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        legend.text = element_text(size = 8.5),
        strip.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.margin = margin(5, 25, 5, 25))+
  coord_cartesian(clip="off")

ggsave("thesis_figures/highEmission_error_pointPlot.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")

```

Mixed baskets

```{r}
# Create label data for the right facet
label_data_mixed <- data.frame(
  x = 2.3,  # adjust to place right of rightmost group
  y = -0.08,
  label = "Perfect accuracy",
  facet = factor("Mixed", levels = c("Other", "Mixed"))
)

# Make sure facet column in your main data matches this
plot_mixed_error <- ggpredict(model_mixedBasket, terms = c("phase", "condition", "mixedBasket"))
plot_mixed_error$facet <- factor(plot_mixed_error$facet, levels = c("Other", "Mixed"))

# Create shaded baseline rectangles for both facets
baseline_rect_mixed <- data.frame(
  facet = factor(c("Other", "Mixed"), levels = c("Other", "Mixed")),
  xmin = 0.5,
  xmax = 1.5,
  ymin = -Inf,
  ymax = Inf
)

# Plot
ggplot(plot_mixed_error, aes(x = x, y = predicted, color = group)) +
  geom_rect(data = baseline_rect_mixed,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey60", alpha = 0.2, inherit.aes = FALSE) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  facet_wrap(~facet, labeller = as_labeller(c("Other" = "Other baskets", "Mixed" = "Mixed baskets"))) +
  scale_color_manual(values = farver, labels = c("No additional support   ",         
                                                 "Semantic support   ",
                                                 "Visuospatial support")) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",
                              keywidth = unit(-5, "cm"),
                              default.unit = "cm")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_label(data = label_data_mixed,
             aes(x = x, y = y, label = label),
             inherit.aes = FALSE, color = "black", fill = "white", label.size = NA, size = 2.6) +
  labs(
    x = "",
    y = "Predicted Directional Error",
    color = "Support condition"
  ) +
  scale_y_continuous(limits = c(-1.1, 1)) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        legend.text = element_text(size = 8.5),
        strip.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.margin = margin(5, 25, 5, 25)) +
  coord_cartesian(clip = "off")

ggsave("thesis_figures/mixed_error_pointPlot.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")

```

mixed without rectangles
```{r}
# Create label data
label_data_mixed <- data.frame(
  x = "Intervention",
  predicted = 0,
  facet = factor("Mixed", levels = c("Other", "Mixed")) # must be a factor with same levels
)

# Predictions
plot_mixed_error <- ggpredict(model_mixedBasket, terms = c("phase", "condition", "mixedBasket"))


# Plot
ggplot(plot_mixed_error, aes(x = x, y = predicted, color = group)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  facet_wrap(~facet, labeller = as_labeller(c("Other" = "Other baskets", "Mixed" = "Mixed baskets"))) +
  scale_color_manual(values = farver, labels = c("No visual support   ",         # add spaces after label
                                                 "Semantic support   ",
                                                 "Visuospatial support")) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           # label right of dot
                              keywidth = unit(-5, "cm"),         # tiny space before label
                              default.unit = "cm"))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_label(data = label_data_mixed, aes(x = max(as.numeric(plot_mixed_error$x)) + 0.28, y = -0.1, label = "Perfect accuracy"), 
           inherit.aes = FALSE, color = "black", fill = "white", label.size = 0, size = 2.5)+
  labs(
    #title = "Mixed-bakset bias",
    x = "",
    y = "Predicted Directional Error",
    color = "Support condition"
  ) +
  scale_y_continuous(limits = c(-1.1, 1)) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        strip.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.margin = margin(5, 25, 5, 25))+
  coord_cartesian(clip="off")

ggsave("thesis_figures/mixed_error_pointPlot.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")
```

Certainty
```{r}
# Predictions
plot_certainty <- ggpredict(certainty_model, terms = c("phase", "condition"))



ggplot(plot_certainty, aes(x = x, y = predicted, color = group)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1, position = position_dodge(0.5)) +
  scale_color_manual(values = farver) +
  labs(
    x = "",
    y = "Predicted certainty",
    color = "Scale condition"
  ) +
  theme_minimal()+
  scale_y_continuous(limits = c(0.5, 7.5)) +
  theme(legend.position = "bottom")
```

Delta certainty
```{r}
# Predictions
plot_delta_certainty <- ggpredict(delta_Certainty_model, terms = c("condition"))

ggplot(plot_delta_certainty, aes(x = group, y = predicted, color = x)) +
  geom_point(position = position_dodge(0.5), size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.1,
                position = position_dodge(0.5)) +
  scale_color_manual(values = farver, labels = c("No additional support   ",         # add spaces after label
                                                 "Semantic support   ",
                                                 "Visuospatial support")) +
  guides(color = guide_legend(override.aes = list(linetype = 0,
                                                  shape = 16,
                                                  size = 3),
                              label.position = "right",           # label right of dot
                              keywidth = unit(-5, "cm"),         # tiny space before label
                              default.unit = "cm"))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  annotate("text", x = 1.2, 
           y = 0.1, label = "No difference between conditions", hjust = 0, size = 2.6, color = "black") + 
  labs(
    #title = "Confidence",
    x = "",
    y = "Predicted Δ Certainty",
    color = "Support condition"
  ) +
  scale_y_continuous(limits = c(-1.5, 1.5)) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  coord_cartesian(clip = "off")

ggsave("thesis_figures/confidence_pointPlot.png", width = 5, height = 4, dpi = 300)

```


# Plots - REPORT

 Vertical bar plot (function)
```{r}
plot_signed_error <- function(data, title_text = NULL, y_limits = c(-2, 2)) {
  
  ggplot(data, aes(x = x, y = predicted, fill = group)) +
    geom_col(position = position_dodge2(0.75, reverse = FALSE, padding = 0.05), width = 0.6) +

    # Annotate reference lines and labels
    annotate("text", x = 0.04, y = 0, hjust = 0, label = "Korrekt\nvurdering",
          family = "Times", fontface = "italic", color = "#034D69", size = 4)+
    annotate("text", x = 0.05, y = 1, hjust = 0, label = "Overvurdering", 
           color = "#034D69", family = "Times", fontface = "italic", size = 4) +
    annotate("text", x = 0.05, y = -1, hjust = 0, label = "Undervurdering", 
           color = "#034D69", family = "Times", fontface = "italic", size = 4)+
    annotate("text", x = 1, y = 0.9, label = "Del 1", fontface = "bold", 
         family = "Times", size = 4.5, hjust = 0.5, color = "#034D69") +
    annotate("text", x = 2, y = 0.9, label = "Del 2", fontface = "bold", 
         family = "Times", size = 4.5, hjust = 0.5, color = "#034D69") +
    annotate("text", x = 1, y = 0.8, label = "Kontrolfase uden støtte", 
         family = "Times", size = 4.5, hjust = 0.5, color = "#034D69") +
    annotate("text", x = 2, y = 0.8, label = "Interventionsfase med støtte",
         family = "Times", size = 4.5, hjust = 0.5, color = "#034D69") +
    annotate("segment", x = -0, xend = Inf, y = 0, yend = 0, 
         linetype = "dashed", colour = "#034D69", linewidth = 0.5)+
  

    # Scales and labels
    scale_fill_manual(values = farver) +
  
    scale_y_continuous(limits = c(-1, 1),
                     breaks = c(-1, 0, 1),
                     labels = c("-1", "0", "+1")) +

    coord_cartesian(clip = "off") +
    labs(
      title = title_text,
      x = "",
      y = expression("Gennemsnitlig afvigelse fra korrekt CO"[2]*"-mærke"),
      fill = "Grad af visuel støtte"
    ) +

    # Theme
    theme_minimal(base_size = 14, base_family = "Times") +
    theme(
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.background = element_blank(),
      legend.position = "none",
      axis.title = element_text(face = "bold", color = "#034D69", vjust = +0.7),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_text(color = "#034D69"),
      axis.line.y = element_line(color = "#034D69", linewidth = 0.6),
      axis.ticks.y = element_line(color = "#034D69"),
      plot.title = element_text(size = 16, face = "bold", color = "#034D69", hjust = 0.5),
      plot.margin = margin(t = 5, r = 10, b = 5, l = 10)
    )
}

# Custom colours
farver <- c(
  "Visuospatial support" = "#e4c358",
  "Semantic support" = "#007A6C",
  "No visual support" = "#034D69"
)

```

Vertical bar plot INTERVENTION ONLY  (function)
```{r}
plot_intervention_signed_error <- function(data, title_text = NULL, y_limits = c(-1, 1)) {
  
  # Behold kun Intervention-fase
  data <- data %>% filter(x == "Intervention")
  
  ggplot(data, aes(x = group, y = predicted, fill = group)) +
    geom_col(position = position_dodge2(0.75, reverse = FALSE, padding = 0.05), width = 0.6) +

    annotate("segment", x = -Inf, xend = Inf, y = 0, yend = 0, 
             linetype = "dashed", colour = "#034D69", linewidth = 0.5) +

    # Skala og labels
    scale_fill_manual(values = farver) +
    scale_y_continuous(limits = y_limits,
                       breaks = c(-1, 0, 1),
                       labels = c("-1", "0", "+1")) +

    coord_cartesian(clip = "off") +
    labs(
      title = title_text,
      x = "",
      y = NULL,  # Fjern y-aksen titel
      fill = "Grad af visuel støtte"
    ) +

    # Tema
    theme_minimal(base_size = 14, base_family = "Times") +
    theme(
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.background = element_blank(),
      legend.position = "none",
      axis.title = element_text(face = "bold", color = "#034D69", vjust = +0.7),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_text(color = "#034D69"),
      axis.line.y = element_line(color = "#034D69", linewidth = 0.6),
      axis.ticks.y = element_line(color = "#034D69"),
      plot.title = element_text(size = 16, face = "bold", color = "#034D69", hjust = 0.5),
      plot.margin = margin(t = 5, r = 10, b = 5, l = 10)
    )
}

```


Directional error
- Only intervention
```{r}
# Get predicted values
pred_signed_error <- ggpredict(signedErrorModel, terms = c("phase", "condition"))

# plot
plot_intervention_signed_error(pred_signed_error)

# save
ggsave("report_figures/signed_error_intervention_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")
```

- with phases
```{r}
# Get predicted values
pred_signed_error <- ggpredict(signedErrorModel, terms = c("phase", "condition"))

# Plot
plot_signed_error(
  data = pred_signed_error,
  title_text = ""
)
# Save
ggsave("report_figures/signed_error_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")

```


High emission
- only intervention
```{r}
plot_highImpact_error <- ggpredict(model_highImpact, terms = c("phase", "condition", "high_impact"))
plot_high <- subset(plot_highImpact_error, facet == "High")

# Plot high emission (D or E)
plot_signed_error(data = plot_high)

# Save
ggsave("report_figures/highEmission_error_intervention_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")


```

- with both phases
```{r}
# Get predictions
plot_highImpact_error <- ggpredict(model_highImpact, terms = c("phase", "condition", "high_impact"))

# Filter  data
plot_high <- subset(plot_highImpact_error, facet == "High")

# Plot high emission (D or E)
plot_signed_error(
  data = plot_high,
  title_text = expression("Kurve med højt gennemsnitligt CO"[2]*"-aftryk")
)

# Save
ggsave("report_figures/highEmission_error_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")

```

Low emission
- with both phases
```{r}
plot_low <- subset(plot_highImpact_error, facet == "Low")
# Plot low emission (A–C)
plot_signed_error(
  data = plot_low,
  title_text = expression("Kurve med lavt gennemsnitligt CO"[2]*"-aftryk")
)

# Save
ggsave("report_figures/lowEmission_error_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")

```


Mixed
- only intervention
```{r}
# predction
plot_mixed_error <- ggpredict(model_mixedBasket, terms = c("phase", "condition", "mixedBasket"))

# subset
plot_mixed <- subset(plot_mixed_error, facet == "Mixed")

# plot
plot_intervention_signed_error(data = plot_mixed )

# Save
ggsave("report_figures/mixed_error_intervention_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")

```

- with both phases
```{r}
# Get predictions
plot_mixed_error <- ggpredict(model_mixedBasket, terms = c("phase", "condition", "mixedBasket"))

# Filter data
plot_mixed <- subset(plot_mixed_error, facet == "Mixed")

# Plot
plot_signed_error(
  data = plot_mixed,
  title_text = expression("Indkøbskurve med både klimavenlige og klimabelastende varer")
)

# Save
ggsave("report_figures/mixed_error_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")
```

Other
```{r}
# Get predictions
plot_mixed_error <- ggpredict(model_mixedBasket, terms = c("phase", "condition", "mixedBasket"))

# Filter data
plot_other <- subset(plot_mixed_error, facet == "Other")

# Plot
plot_signed_error(
  data = plot_other,
  title_text = expression("Indkøbskurve med varer uden stor forskel i klimaaftryk")
)

# Save
ggsave("report_figures/other_error_plot.png", width = 6.5, height = 4.5, dpi = 300, bg = "transparent")
```

Delta certainty - horizontal bar
```{r}

plot_delta_certainty <- ggpredict(delta_Certainty_model, terms = c( "condition"))


ggplot(plot_delta_certainty, aes(y = x, x = predicted, fill = x)) +
  geom_col(width = 0.6) +
  #geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_fill_manual(values = farver, guide = guide_legend(reverse = FALSE)) +
  
  # Annotations
  annotate("segment", y = -Inf, yend = Inf, x = 0, xend = 0,
           linetype = "dashed", colour = "#034D69", linewidth = 0.5) +
  
  labs(
    y = "",
    x = "",
    fill = ""
  ) +
  
  scale_x_continuous(limits = c(-1.2, 1.2),  position = "top")+#, breaks = seq(-1.5, 0.5, 0.5), expand = c(0, 0)) +
  scale_y_discrete(labels = c(
    "No visual support" = "Ingen visuel støtte",
    "Semantic support" = "Semantisk støtte",
    "Visuospatial support" = "Rummelig støtte"))+
  # Theme
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    legend.position = "none",
    axis.title.x = element_text(face = "bold", color = "#034D69"),
    axis.text.x = element_text(color = "#034D69"),
    axis.text.y = element_blank(),
    axis.line.x = element_line(color = "#034D69", linewidth = 0.6),
    axis.ticks.x = element_line(color = "#034D69"),
    plot.margin = margin(t = 5, r = 5, b = 5, l = -5)
  )

# Save
ggsave("report_figures/certainty_vertical_plot.png", width = 6, height = 4.5, dpi = 300, bg = "transparent")

```


# Assumptions

Absolute error
```{r}
# Extract residuals and fitted values
resid_abs <- resid(absErrorModel)
fitted_abs <- fitted(absErrorModel)

# Histogram of residuals
hist(resid_abs, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_abs)
qqline(resid_abs)

# Residuals vs. fitted
plot(fitted_abs, resid_abs,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

Directional error
```{r}
# Extract residuals and fitted values
resid_signed <- resid(signedErrorModel)
fitted_signed <- fitted(signedErrorModel)

# Histogram of residuals
hist(resid_signed, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_signed)
qqline(resid_signed)

# Residuals vs. fitted
plot(fitted_signed, resid_signed,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

High Emission
```{r}
# Extract residuals and fitted values
resid_high <- resid(model_highImpact)
fitted_high <- fitted(model_highImpact)

# Histogram of residuals
hist(resid_high, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_high)
qqline(resid_high)

# Residuals vs. fitted
plot(fitted_high, resid_high,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)

```

Mixed
```{r}
# Extract residuals and fitted values
resid_mixed <- resid(model_mixedBasket)
fitted_mixed <- fitted(model_mixedBasket)

# Histogram of residuals
hist(resid_mixed, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_mixed)
qqline(resid_mixed)

# Residuals vs. fitted
plot(fitted_mixed, resid_mixed,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

Certainty
```{r}
# Extract residuals and fitted values
resid_certainty <- resid(delta_Certainty_model)
fitted_certainty <- fitted(delta_Certainty_model)

# Histogram of residuals
hist(resid_certainty, breaks = 50, main = "Residuals Histogram", xlab = "Residuals")

# QQ plot
qqnorm(resid_certainty)
qqline(resid_certainty)

# Residuals vs. fitted
plot(fitted_certainty, resid_certainty,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```


# Demographics - plots for the report

Data
```{r}
df_unique <- df %>%
  distinct(participantId, gender, age, occupation, income)

# Define theme
my_theme <- theme_minimal(base_family = "Times", base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", color = "#034D69", hjust = 0.5),
    axis.text = element_text(color = "#034D69", face = "bold", size = 12),
    axis.title = element_text(color = "#034D69", face = "bold"),
    axis.text.y = element_text(margin = margin(r = -12)),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    panel.background = element_blank()
  )

# Color
blue <- "#034D69"
```

Gender
```{r}
df_unique %>%
  count(gender) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = "", y = prop, fill = gender)) +
  geom_col(width = 1, color = "white") +
  coord_polar("y") +
  geom_text(aes(label = paste0(round(prop*100), "%")), 
            position = position_stack(vjust = 0.5), color = "white", size = 5) +
  scale_fill_manual(values = c("Kvinde" = "#e4c358", "Mand" = "#007A6C", "Andet" = "#034D69")) +
  labs(title = "Fordeling på køn", fill = NULL) +
  my_theme +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank())


df_unique %>%
  count(gender) %>%
  mutate(percent = round(n / sum(n) * 100)) %>%
  ggplot(aes(x = fct_rev(gender), y = n)) +
  geom_col(fill = blue, width = 0.5) +
  geom_text(aes(label = paste0(percent, "%")), hjust = -0.1, color = "#034D69", fontface = "bold", family = "times", size = 4.5) +
  scale_fill_manual(values = gender_colors) +
  coord_flip(clip = "off") +
  labs(title = "Fordeling på køn", x = NULL, y = NULL) +
  my_theme +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 5, r = 10, b = 5, l = 5)
  )

gender_colors <- c(
  "Kvinde" = "#689aae",  # light blue (far left)
  "Mand" = "#84bf7c",    # dark blue (second from left)
  "Andet" = "#095979"    # very light mint/blue (second from right)
)

df_unique %>%
  count(gender) %>%
  mutate(
    percent = round(n / sum(n) * 100),
    label = paste0(gender, "\n", percent, "%"),
    angle = 90 - 360 * (cumsum(n) - 0.5 * n) / sum(n)
  ) %>%
  ggplot(aes(x = 2, y = n, fill = gender)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  #geom_text(aes(label = label, y = n / 2 + c(0, cumsum(n)[-length(n)])),
         #   color = "#034D69", family = "times", fontface = "bold", size = 4.5) +
  scale_fill_manual(values = gender_colors) +
  xlim(0.5, 2.5) +
  theme_void() +
  labs(title = "Fordeling på køn") +
  theme(
    plot.title = element_text(hjust = 0.5, family = "times", face = "bold", colour = "#034D69"),
    legend.position = "none",
    plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
  )

ggsave("demografi_køn_donutPlot.png", width = 6, height = 6, dpi = 300, bg = "transparent")



```

Age groups
```{r}
df_unique <- df_unique%>%
  mutate(age_group = cut(age, 
                         breaks = c(17, 29, 39, 49, 59, 69, Inf), 
                         labels = c("18-29 år", "30-39 år", "40-49 år", 
                                    "50-59 år", "60-69 år", "70+ år")))
df_unique <- df_unique %>%
  mutate(age_group = factor(age_group, levels = c(
    "18-29 år", "30-39 år", "40-49 år",
    "50-59 år", "60-69 år", "70+ år"
  )))

df_unique %>%
  count(age_group) %>%
  mutate(percent = round(n / sum(n) * 100)) %>%
  ggplot(aes(x = age_group, y = n)) +
  geom_col(fill = blue, width = 0.5) +
  geom_text(aes(label = paste0(percent, "%")),
            hjust = -0.2, color = blue, fontface = "bold", family = "times", size = 4.5, clip = "off") +
  coord_flip(clip = "off") +
  labs(title = "Aldersgrupper", x = NULL, y = NULL) +
  my_theme +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 5, r = 10, b = 5, l = 5)
  )



ggsave("demografi_alder_plot.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")


```

Occupation
```{r}
df_unique %>%
  count(occupation) %>%
  mutate(
    percent = round(n / sum(n) * 100),
    occupation = fct_rev(fct_reorder(occupation, n))
  ) %>%
  ggplot(aes(x = occupation, y = n)) +
  geom_col(fill = blue, width = 0.5) +
  geom_text(aes(label = paste0(percent, "%")), 
            hjust = -0.4, color = blue, fontface = "bold", family = "Times", size = 4.5) +
  labs(title = "Beskæftigelse", x = NULL, y = NULL) +
  coord_flip(clip = "off") +
  my_theme +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 5, r = 40, b = 5, l = 5)
  )

ggsave("demografi_beskaeftigelse_plot.png", width = 6.5, height = 3, dpi = 300, bg = "transparent")

```

Income
```{r}

income_levels <- c(
  "Under 10.000",
  "10–25.000",
  "25–40.000",
  "40–55.000",
  "55–70.000",
  "Mere end 70.000",
  "Ønsker ikke at oplyse"
)


df_unique %>%
  count(income) %>%
  mutate(
    percent = round(n / sum(n) * 100),
    income = factor(income, levels = income_levels)
  ) %>%
  ggplot(aes(x = income, y = n)) +
  geom_col(fill = blue, width = 0.5) +
  geom_text(aes(label = paste0(percent, "%")),
            hjust = -0.2, color = blue, fontface = "bold", size = 4.5, family = "Times") +
  coord_flip(clip = "off") +
  labs(title = "Indkomstfordeling", x = NULL, y = NULL) +
  my_theme +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 5, r = 40, b = 5, l = 5)
  )

ggsave("demografi_indkomst_plot.png", width = 6.5, height = 4, dpi = 300, bg = "transparent")

```



# National scale estimate - calculations of misjudged kg co2 per misjudged high-emission basket
```{r}
df_high = df %>% 
  filter(high_impact == "High")
```

```{r}
df_high_incorrect = df_high %>% 
  filter(error > 0)
```

```{r}
# label midpoint mapping based on intervals
label_midpoints <- c(
  A = 0.55,
  B = 1.5,
  C = 3.75,
  D = 7.5,
  E = 20.5  # Precise average for ten E products used in experiment
)

# Create column with judged CO₂ based on the midpoint of selectedLabel
df_high_incorrect$guessedCo2 <- label_midpoints[df_high_incorrect$selectedLabel]

# Calculate underestimation
df_high_incorrect$underestimation <- df_high_incorrect$weightedCo2 - df_high_incorrect$guessedCo2
```

```{r}
# Summary statistics
mean_judgement_error <- mean(df_high_incorrect$underestimation, na.rm = TRUE)
total_judgement_error <- sum(df_high_incorrect$underestimation, na.rm = TRUE)

underestimates <- df_high_incorrect[df_high_incorrect$underestimation > 0, ]
n_underestimated <- nrow(underestimates)



mean_underestimation <- mean(underestimates$underestimation, na.rm = TRUE)
total_underestimation <- sum(underestimates$underestimation, na.rm = TRUE)

cat("Mean misjudgement in co2 per basket:", round(mean_judgement_error, 2), "kg\n")
cat("Total misjudgement in co2 across", n_underestimated, "trials:", round(total_judgement_error, 2), "kg\n")

cat("Mean underestimation per misjudged basket:", round(mean_underestimation, 2), "kg\n")
cat("Total underestimation across", n_underestimated, "trials:", round(total_underestimation, 2), "kg\n")

```